---
title: "Problem_2"
author: "Shumin Rui"
date: "10/24/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = 1,
  out.width = "120%")
library(readxl)
library(tidyverse)
library(dplyr)
library(janitor)
library(lubridate)
```

```{r tidy the data}
#import GDP data
gdp_data = read_csv("./GDP.csv") %>% 
  clean_names() %>% 
  filter(!(value == ".")) %>% 
  mutate(date = as.Date(date))

#import pols dataset
month_data = read_csv("./pols-month.csv") %>% 
  clean_names() %>% 
  na.omit() %>% 
  rename(date = mon) %>% 
  mutate(date = floor_date(ymd(date), "month")) %>% 
  mutate(president = recode(prez_gop, `0` = "dem", `1` = "gop", `2` = "gop")) %>%
  select(-prez_gop, -prez_dem)

#combine two datasets
gdp_mon_data <- left_join(gdp_data, month_data, by = "date")
```


```{r plot_GDP}
gdp_mon_data %>%
  ggplot( aes(x = as.character(date), y = as.numeric(value), fill = president)) + # plot GDP change over time, color of bins depends on president variable
  geom_bar(stat="identity", width = 1) + # change binwidth to 1
  labs(
    title = "GDP change over time", #plot title
    x = "Date", # x label
    y = "GDP" # y label
  ) +
  theme(legend.position = "bottom", # put legend to the bottom
        text = element_text(size = 12), # set test front size
        axis.text.x = element_text(size = 4, angle = 90, hjust = 1) # set x asix test style
        )
```

The plot above shows the quarterly GDP change over time from `r gdp_mon_data[1, "date"]` to `r gdp_mon_data[nrow(gdp_mon_data), "date"]`. The dates when the president was democrat are labeled with red, and when the president was republican are labeled in blue. According to the plot, it is hard to find the association between GDP and the political party of the president, but we can see how it change over time.
